---
title: "Cyrius Post-Run Report"
author:
  - "Lightning Auriga"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  rmd: "cyrius.Rmd"
output:
  html_document:
    code_folding: "hide"
    highlight: tango
    number_sections: no
    theme: default
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r load.packages, eval=TRUE, echo=FALSE}
## Load required R packages

require(knitr, quietly = TRUE)
require(kableExtra, quietly = TRUE)
require(ggplot2, quietly = TRUE)
require(RColorBrewer, quietly = TRUE)
require(stringr, quietly = TRUE)
require(DescTools, quietly = TRUE)
```

```{r set.ggplot2.theme, eval=TRUE, echo=TRUE}
## Configure standard theme information for ggplot2

my.theme <- theme_light() + theme(
  plot.title = element_text(size = 15, hjust = 0.5),
  axis.title = element_text(size = 14),
  axis.text = element_text(size = 12),
  strip.background = element_blank(),
  strip.text = element_text(size = 14, colour = "black")
)
```

```{r set.variables, eval=TRUE, echo=TRUE}
## Set input variables from snakemake
cyrius.tsv <- snakemake@input[["results"]]
cyrius.published.table <- snakemake@input[["pubtable"]]
somalier.ancestry.tsv <- snakemake@input[["somalier"]]
```

```{r sanitize.asterisks.function, eval=TRUE, echo=FALSE}
sanitize.asterisks <- function(vec) {
  stringr::str_replace_all(vec, "\\*", "\\\\*")
}
```

```{r load.somalier.ancestry, eval=TRUE, echo=FALSE}
somalier.ancestry <- read.table(somalier.ancestry.tsv,
  header = FALSE,
  stringsAsFactors = FALSE, sep = "\t",
  comment.char = "#", quote = "",
  check.names = FALSE
)[, 1:2]
colnames(somalier.ancestry) <- c("id", "ancestry")
somalier.ancestry$id <- stringr::str_split_i(somalier.ancestry$id, "_", 1)
## eventually, handle or warn about conflicting duplicates
somalier.ancestry <- somalier.ancestry[!duplicated(somalier.ancestry$id), ]
rownames(somalier.ancestry) <- somalier.ancestry$id
```

```{r load.cyrius.table, eval=TRUE, echo=TRUE}
cyrius.data <- read.table(cyrius.tsv, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
## only permit subjects who have ancestry estimates from somalier; this normally shouldn't
## be an issue, but good to be sure.
cyrius.data <- cyrius.data[cyrius.data[, 1] %in% somalier.ancestry[, 1], ]
cyrius.table.data <- cyrius.data
cyrius.table.data[, 2] <- sanitize.asterisks(cyrius.table.data[, 2])
na12878.present <- length(which(cyrius.table.data[, 1] == "NA12878")) > 0
```


```{r initialize.known.counts, eval=TRUE, echo=FALSE}
cyrius.published.data <- read.table(cyrius.published.table,
  header = TRUE,
  sep = "\t", stringsAsFactors = FALSE,
  check.names = FALSE
)
cyrius.published.data <- cyrius.published.data[, 1:7]
```

## Haplotype Call Summary Metrics

```{r report.basic.summary.metrics, eval=TRUE, echo=FALSE, results="asis"}
compute.metric.data <- function(cyrius.data) {
  calls <- c(
    "PASS",
    "More_than_one_possible_genotype",
    "Not_assigned_to_haplotypes",
    "LowQ_high_CN",
    "None",
    "[combined]"
  )
  counts <- c(
    length(which(cyrius.data[, 3] == calls[1])),
    length(which(cyrius.data[, 3] == calls[2])),
    length(which(cyrius.data[, 3] == calls[3])),
    length(which(cyrius.data[, 3] == calls[4])),
    length(which(cyrius.data[, 3] == calls[5])),
    nrow(cyrius.data)
  )
  metric.data <- data.frame(
    Call = calls,
    Count = counts
  )
  metric.data
}

metric.data <- compute.metric.data(cyrius.data)
knitr::kable(metric.data, caption = "Haplotype Call Summary Metrics") %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
```

***
<br>

`r if (na12878.present) {"## Positive Control Report"}`

`r if (na12878.present) {cat("Per [this paper](https://www.nature.com/articles/s41397-020-00205-5),",
                             "the expected haplotype call for NA12878 is \\*3/\\*68+\\*4.")}`

```{r positive.control.report, eval=na12878.present, echo=FALSE, results="asis"}
knitr::kable(cyrius.table.data[cyrius.table.data[, 1] == "NA12878", ],
  caption = "Positive Control Results for Cyrius"
) %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
cat("\n\n***\n\n<br>\n\n")
```


## Haplotype Frequencies versus 1KG

The following table recapitulates the contents of [Table 2 of the Cyrius paper](https://doi.org/10.1038/s41397-020-00205-5),
adding an additional column representing the corresponding haplotype frequency data from the current subject set.

```{r annotate.counts.from.dataset, eval=TRUE, echo=FALSE}
compute.pub.data.table <- function(cyrius.data, cyrius.published.data, somalier.ancestry, target.ancestry) {
  out.df <- cyrius.published.data
  out.df$"This dataset" <- rep(0, nrow(out.df))
  for (i in seq_len(nrow(cyrius.data))) {
    if (somalier.ancestry[cyrius.data[i, 1], "ancestry"] != target.ancestry) {
      next
    }
    if (cyrius.data[i, 3] == "PASS") {
      gtype1 <- stringr::str_replace(cyrius.data[i, 2], "^([^/]+)/.*$", "\\1")
      gtype2 <- stringr::str_replace(cyrius.data[i, 2], "^[^/]+/(.*)$", "\\1")
      update.count <- function(gtype) {
        res <- out.df$"This dataset"
        if (length(which(out.df[, 1] == gtype)) == 1) {
          res[out.df[, 1] == gtype] <-
            out.df$"This dataset"[out.df[, 1] == gtype] + 1
        } else {
          res[out.df[, 1] == "Unknown"] <-
            out.df$"This dataset"[out.df[, 1] == "Unknown"] + 1
        }
        res
      }
      out.df$"This dataset" <- update.count(gtype1)
      out.df$"This dataset" <- update.count(gtype2)
    } # else if (cyrius.data[i, 3] == "More_than_one_possible_genotype") {
    ##
    ## } else if (cyrius.data[i, 3] == "Not_assigned_to_haplotypes") {
    ##
    ## } # else cyrius.data[i, 3] == "LowQ_high_CN"
  }

  out.df[, 1] <- sanitize.asterisks(out.df[, 1])
  multi.cis <- DescTools::MultinomCI(out.df$"This dataset")
  out.df$"This dataset" <- round(multi.cis[, 1] * 100, 2)
  out.df$"95% CI" <- paste("[", round(multi.cis[, 2] * 100, 2), ", ",
    round(multi.cis[, 3] * 100, 2), "]",
    sep = ""
  )
  out.df$"This dataset"[nrow(out.df)] <- NA
  out.df
}

anc.linker <- c(
  "Admixed American" = "AMR",
  "East Asian" = "EAS",
  "African" = "AFR",
  "South Asian" = "SAS",
  "European" = "EUR"
)

prune.ancestry.columns <- function(df, anc) {
  if (anc %in% anc.linker) {
    possible.colnames <- colnames(df)[3:7]
    for (possible.colname in possible.colnames) {
      if (anc.linker[possible.colname] != anc) {
        df[, possible.colname] <- NULL
      }
    }
  }
  df
}
```

```{r report.counts.from.dataset, eval=TRUE, echo=FALSE, results="asis"}
target.ancestries <- sort(unique(somalier.ancestry[cyrius.data[, 1], "ancestry"]))

for (target.ancestry in target.ancestries) {
  cat("### ", names(anc.linker)[which(anc.linker == target.ancestry)], "\n\n", sep = "")
  table.df <- compute.pub.data.table(cyrius.data, cyrius.published.data, somalier.ancestry, target.ancestry)
  table.df <- prune.ancestry.columns(table.df, target.ancestry)
  print(knitr::kable(table.df,
    caption = paste("Cyrius Paper Table 2, with Current ", target.ancestry,
      " Results, as Percentage of Observed Genotypes",
      sep = ""
    )
  ) %>%
    kableExtra::kable_styling("condensed", position = "left", full_width = FALSE))
  cat("\n\n***\n\n<br>\n\n")
}
```

```{r paired.barplot, eval=TRUE, echo=FALSE}
```

## Full Haplotyping Results for All Subjects

```{r report.summary.table, eval=TRUE, echo=FALSE, results="asis"}
cyrius.table.data$Ancestry <- somalier.ancestry[cyrius.table.data[, 1], "ancestry"]
rownames(cyrius.table.data) <- NULL
knitr::kable(cyrius.table.data, caption = "Current Results from Cyrius, with Somalier Ancestry") %>%
  kableExtra::kable_styling("condensed", position = "left", full_width = FALSE)
```


***
<br>

## Assorted Software Links

<br>

- [Cyrius](https://github.com/Illumina/Cyrius)

***
<br>

## Session Information

<br>

The following summarizes the loaded R configuration for the run that created this report.

```{r session.info, eval=TRUE, echo=TRUE}
sessionInfo()
```
